---
title: "RepData_PeerAssessment1"
author: "Amanda Molling"
date: "Saturday, March 14, 2015"
output: html_document
---
This is the summary of the data that was read in:   

```{r}
unzip("activity.zip")
 activityData <- read.csv("activity.csv")
summary(activityData)
head(activityData)
```
If I wanted to convert my dates to type date, I could use. 

```{r, echo=TRUE}
activityData$date <- as.Date(activityData$date)


```
I computed the mean, median, and sum data as follows:

```{r}
totalDailySteps <- aggregate(steps ~ date, activityData, sum)
meanDailySteps <- aggregate(steps ~ date, activityData, mean)
medDailySteps <- aggregate(steps ~ date, activityData, mean)

head(totalDailySteps)
```
Then my histogram:

```{r}
hist(totalDailySteps[, 2], breaks = 20, xlab = "Total Steps Each Day", main = "Histogram of Total Daily Steps Taken")
```
Followed by analysis on that data:
```{r}
mean(totalDailySteps[, 2])
median(totalDailySteps[,2])
```
Further analysis comparing the average number of steps at each time interval across the days yields:
```{r}
meanDailyStepsByInterval <- aggregate(steps ~ interval, activityData, mean)
library('lattice')
xyplot(steps ~ interval, meanDailyStepsByInterval, xlab = "
       ")
```
The time interval with the most steps is found to be 835.
```{r}

which(meanDailyStepsByInterval == max(meanDailyStepsByInterval[2]), arr.ind = TRUE)
```
where the value is:

```{r}

max(meanDailyStepsByInterval[,2])

## And the specific interval is :

meanDailyStepsByInterval[104, 1]
```

For the second part, I find the number of complete cases and build a table. Then, I can subtract the good cases from the total number of obs. Another way to do this is by simply calculating the number of bad cases as shown.

```{r}
##I make the good data first
good <-complete.cases(activityData)

goodData<-activityData[good,][,]
nrow(goodData)
nrow(activityData) - nrow(goodData)

##That is the number of bad rows
## The second method:
badData <- activityData[!good,][,]
nrow(badData)
```
# Replacing Missing Data
### To fill in the missing data, I use my mean data calcuated at each time interval from before:
```{r}
newActivityData <- activityData
for (i in 1 : (nrow(activityData) )){
  if (  is.na( activityData[ i , 1] ) ) {
    newActivityData[ i , 1] <- meanDailyStepsByInterval[ ( ( i - 1 ) %% ( nrow(meanDailyStepsByInterval) ) + 1 ), 2]
  }
}
```
### As confirmation of my replacement, I have shown the front and back of each of the tables, where NA values dominate. 

```{r}
head(newActivityData)
tail(newActivityData)
head(meanDailyStepsByInterval)
tail(meanDailyStepsByInterval)
```

##The Histogram of the Data with Replaced Values:
```{r}
totalDailySteps <- aggregate(steps ~ date, newActivityData, sum)
hist(totalDailySteps[, 2], breaks = 20, xlab = "Total Steps Each Day", main = "Histogram of Total Daily Steps Taken")
```

## The mean and median of the data with artificially replaced  values are:
```{r}

mean(totalDailySteps[,2])
median(totalDailySteps[,2])
```
By looking the mean (107.66.19) and median (10765) from before, it can be seem that any change in the mean is not detectable. In addition, due to relative proximity of the median and mode, the change in the median was small and calculated to be less than 2 hundreths of a percent change:

```{r}
percentDifference <- 100 * (10766.19 - 10765)/(10765)
print(percentDifference)
```
